---
title: "Analises Modelos Mistos"
author: "LucianoRogerio e HenriqueBernardino"
date: "2021-10-12"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

[Wrong Page?](index.html)

## Mixed Models Foliar Disease

```{r MixedModels and Anova Foliar Disease traits}
library(here)
suppressMessages(library(tidyverse))
suppressMessages(library(plyr))
library(reactable)
suppressMessages(library(data.table))
suppressMessages(source(here::here("code", "MixedModelsFunctions.R")))

PhenoData <- readRDS(here::here("output", "DadosDoencasv2.RDS"))

PhenoData$block_number <- as.character(PhenoData$block_number)
PhenoData2 <- PhenoData %>% filter(!is.na(Y)) %>% 
  mutate(TrialBlock = paste0(trial_name, block_number))
traits <- unique(PhenoData2$traits)

fmfit <- PhenoData2 %>% dlply(.variables = c("traits"),
                              .fun = analyzeTrial.lme4FD)
ResFixEffect <- lapply(fmfit, FUN = as.data.frame(anova))
ResAnInt <- matrix(unlist(ResFixEffect,use.names = T),
                   nrow = 2, byrow = F)
ResAnFin <- rbind(ResAnInt[,1:4],
                  ResAnInt[,5:8],
                  ResAnInt[,9:12],
                  ResAnInt[,13:16])
colnames(ResAnFin) <- c("DF", "SumSq", "MeanSq", "Fvalue")
ResAnovaFinal <- data.frame(Trait = rep(traits, each = 2),
                            Factor = rep(c("Trial", "Trial:Block"),
                                         times = 4),
                            ResAnFin)
```

#### Table 1. Anova of the fixed effects of Cassava foliar diseases
```{r table1, echo = FALSE}
ResAnovaFinal %>% reactable(columns = list(
  SumSq = colDef(format = colFormat(digits = 3, locales = "en-US")),
  MeanSq = colDef(format = colFormat(digits = 3, locales = "en-US")),
  Fvalue = colDef(format = colFormat(digits = 3, locales = "en-US"))))
```


```{r Deviance analysis}
rdfmfit1 <- PhenoData2 %>% dlply(.variables = c("traits"),
                                .fun = analyzeTrialrdMod1.lme4)
rdfmfit2 <- PhenoData2 %>% dlply(.variables = c("traits"),
                                .fun = analyzeTrialrdMod2.lme4)

Deviances <- NULL
for(i in traits){
Deviances[[i]] <- data.frame(Deviance.MM(fmfit[[i]], rdfmfit1[[i]], rdfmfit2[[i]]))[2:3,6:8]
rownames(Deviances[[i]]) <- c("Clone", "GxE")
}
ResDeviances <- matrix(unlist(Deviances, use.names = T), nrow = 2, byrow = F)
ResDeviances <- rbind(ResDeviances[,1:3],
                  ResDeviances[,4:6],
                  ResDeviances[,7:9],
                  ResDeviances[,10:12])
colnames(ResDeviances) <- c("QuiSq", "DF", "Prob")

ResDeviancesFinal <- data.frame(Trait = rep(traits, each = 2),
                            Factor = rep(c("Clone", "GxE"),
                                         times = 4),
                            ResDeviances)
```

#### Table 2. Deviance Analysis for cassava foliar disease
```{r table2, echo = FALSE}
ResDeviancesFinal %>% reactable()
```

```{r heritability estimation}
H2 <- sapply(fmfit, FUN = getVarComp.lme4) %>% t() %>% as.data.frame()
colnames(H2) <- c("VarClone", "VarGE", "VarRes")

MediasFix <- as.matrix(sapply(fmfit, FUN = (fixef))) %>%
  .[rownames(.) == "(Intercept)"] %>% data.frame(Mean = .)

RepMean <- tibble()

for(i in traits){
    RepMean <- bind_rows(RepMean, tibble(Trait = i,
      RepMean = PhenoData2 %>% group_by(traits, trial_name, accession_name) %>%
      dplyr::summarise(RepMean = n())%>% .$RepMean %>% mean()))
}
H2 <- cbind(H2, MediasFix, RepMean)


H2 <- H2 %>% mutate(VarClone = as.numeric(VarClone),
                    VarGE = as.numeric(VarGE),
                    VarRes = as.numeric(VarRes),
                    VarFen = VarClone + VarGE + VarRes,
                    VarFenFam = VarClone + VarGE/RepMean + VarRes/(RepMean*2),
                    H2 = VarClone/VarFen,
                    H2fam = VarClone/VarFenFam,
                    CVg = sqrt(VarClone)/Mean,
                    CVe = sqrt(VarRes)/Mean)
H2[,"Mean"] <- NULL
H2[,"RepMean"] <- NULL
```

#### Table 3. Heritabilities of cassava foliar disease
```{r table3, echo = FALSE}
H2 %>% reactable(columns = list(
  VarClone = colDef(format = colFormat(digits = 4, locales = "en-US")),
  VarRes = colDef(format = colFormat(digits = 4, locales = "en-US")),
  VarFen = colDef(format = colFormat(digits = 4, locales = "en-US")),
  H2 = colDef(format = colFormat(digits = 4, locales = "en-US")),
  H2fam = colDef(format = colFormat(digits = 4, locales = "en-US")),
  CVg = colDef(format = colFormat(digits = 4, locales = "en-US", percent = T)),
  CVe = colDef(format = colFormat(digits = 4, locales = "en-US", percent = T))))
```

## Obtain the Mean + BLUPs of the clones

### Obtain the Clones BLUPS

```{r Estimating the BLUPs + Mean for the clones for foliar disease}
BLUPsAle <- lapply(fmfit, FUN = getBLUPs.lme4)

BLUPSDisea <- data.frame(CLONE = rownames(BLUPsAle[1]$Anth))

for(i in names(BLUPsAle)){
  drg<-data.frame(CLONE = rownames(BLUPsAle[[i]]), stringsAsFactors=F)
  drg[,i] <-BLUPsAle[[i]]
  BLUPSDisea<-merge(BLUPSDisea,drg,by="CLONE",all.x=T)
}

saveRDS(BLUPSDisea, here::here("output", "BLUPsDisease.RDS"))
```

#### Table 4. Blups of the accessions for cassava foliar diseases
```{r table4, echo = FALSE}
BLUPSDisea %>% reactable(defaultPageSize = 25, columns = list(
  Anth = colDef(format = colFormat(digits = 4, locales = "en-US")),
  BlLS = colDef(format = colFormat(digits = 4, locales = "en-US")),
  BrLS = colDef(format = colFormat(digits = 4, locales = "en-US")),
  WhLS = colDef(format = colFormat(digits = 4, locales = "en-US"))))
```

#### Fig 1. Distribution of the BLUPs estimated for cassava foliar diseases
```{r Fig1, echo = FALSE, fig.height = 4, fig.width = 3}
#tiff(here::here("output", "Fig4_DenplotsBLUPsCassavaFoliarDiseaseTraits.tiff"), res = 350,
#     compression = "lzw", units = "cm", width = 20, height = 15)
BLUPSDisea %>% reshape2::melt(id.vars = "CLONE",
                         variable.name = "Trait",
                         value.name = "Value") %>%
  ggplot(aes(x = Value)) + geom_density(adjust = 2.5, fill = "darkgreen") + 
  facet_wrap(~ Trait) +
  theme(legend.position="none",
        panel.spacing = unit(0.1, "lines"),
        axis.ticks.x=element_blank()) + theme_bw()
#dev.off()
```

## Mixed Models

### Yield Traits

```{r reading the yield traits}
suppressMessages(library(lme4)); suppressMessages(library(tidyverse))
library(reactable); library(here)

AgroData <- readRDS(file = here::here("output", "DadosFenotipicos.rds"))

head(AgroData)
```

#### Table 5. Data entry to perform the mxed model analysis for Agronomic Traits
```{r table5, echo = FALSE}
AgroData %>% reactable(defaultPageSize = 25, columns = list(
  y = colDef(format = colFormat(digits = 2, locales = "en-US"))))
```

#### Mixed Models analysis

```{r estimation of heritability and reliability per trial and trait}
Trials <- unique(AgroData$trial)
Results <- tibble()

for(i in Trials){
  traits <- AgroData %>% filter(trial %in% i) %>% .$Trait %>% unique %>% as.character
  results <- tibble()

  for(j in traits) {
  try(MixedModels <- analyzeTrial.lme4(AgroData %>% filter(trial %in% i & Trait %in% j)))
  try(result <- tibble(Trial = i,
                       Trait = j,
                       VarG = as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% .[1,2],
                       VarE = as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% .[2,2],
                       H2 = VarG/(VarG + VarE),
                       Real = suppressWarnings(MuMIn::r.squaredGLMM(MixedModels)[2])))
  try(results <- rbind(results, result))
  rm(MixedModels); rm(result)
  }
  
  Results <- rbind(Results, results)
  rm(traits); rm(results)
}

```

#### Yield data selection

```{r Yield data selection, echo = FALSE}
library(reactable)
TrialsList <- unique(AgroData[,c("Ano", "Campo", "Local", "trial")])

Results2 <- Results %>% right_join(TrialsList, by = c("Trial" = "trial")) %>%
  dplyr::select(Trial, Ano, Campo, Local, everything()) %>% mutate(Selecionado = ifelse(Real > 0.25 & H2 > 0.15, "Sim", "Nao"))


AgroDataSel <- AgroData %>% 
  .[.$trial %in% (Results2 %>% filter(Selecionado == "Sim") %>% .$Trial),]

saveRDS(object = AgroDataSel, file = here::here("data", "DadosFenotipicosSel.rds"))
```


#### Table 6. Heritability and reliability of all trials for yield traits
```{r table6, echo = FALSE}
Results2 %>% reactable(groupBy = c("Trait"), columns = list(
  VarG = colDef(format = colFormat(digits = 2, locales = "en-US")),
  VarE = colDef(format = colFormat(digits = 2, locales = "en-US")),
  H2 = colDef(format = colFormat(digits = 3, locales = "en-US")),
  Real = colDef(format = colFormat(digits = 3, locales = "en-US"))))
```

## Estimação BLUPS e obtenção de médias corrigidas

```{r estimating Blups for yield traits, warning = FALSE}
library(here)
library(furrr)
library(tidyverse)
source(here::here("code", "MixedModelsFunctions.R"))

AgroDataSel <- readRDS(here::here("data", "DadosFenotipicosSel.rds")) %>%
  mutate(trial = as.character(trial),
         rep = as.character(rep),
         repTrial = as.factor(paste(trial, rep, sep =":")))


#NCT <- 4

#plan(sequential)
#RhpcBLASctl::blas_set_num_threads(NCT)

traits <- table(AgroDataSel$Trait) %>% .[order(.)] %>% names

for(i in traits){
  print(paste("Trait", i, sep = " "))
  DataMM <- AgroDataSel %>% filter(Trait == i)
  MM <- analyzeTrial.lme4Conj(DataMM)
  blups <- ranef(MM)$clone + fixef(MM)[1]
  Blups <- tibble(id = rownames(blups),
                  blups = blups$`(Intercept)`)
  colnames(Blups)[2] <- i
  file <- here::here("output", "MixedModels",
                     paste("Blups_", i, ".rds", sep = ""))
  saveRDS(object = Blups, file = file)
  rm(DataMM); rm(MM); rm(blups); rm(Blups); rm(file)
}

BlupsTraits <- readRDS(here::here("output", "MixedModels", "Blups_Vigor12M.rds"))
IDClones <- tibble(id = unique(AgroDataSel$clone) %>% .[order(.)])
BlupsTraits <- IDClones %>% left_join(BlupsTraits, by = "id")

for(i in traits[-1]){
  filename <- paste("Blups_", i, ".rds", sep = "")
  BlupsTraits <- BlupsTraits %>%
    left_join(readRDS(here::here("output", "MixedModels", filename)))
  colnames(BlupsTraits)[colnames(BlupsTraits) == "blups"] <- i
}

saveRDS(object = BlupsTraits,
        file = here::here("output", "BlupsFenHen.rds"))
```

#### Fig 2. Distribution of the BLUPs estimated for cassava agronomic traits
```{r Fig2, echo = FALSE, fig.height = 4, fig.width = 3}
#tiff(filename = here::here("output", "Fig5_DenplotsBLUPsCassavaAgronomicTraits.tif"), res = 350,
#     compression = "lzw", units = "cm", width = 20, height = 15)
BlupsTraits %>% dplyr::rename(Vigor = Vigor12M) %>% reshape2::melt(id.vars = "id",
                         variable.name = "Trait",
                         value.name = "Value") %>%
  ggplot(aes(x = Value)) + geom_density(adjust = 2.5, fill = "darkgreen") + 
  facet_wrap(~ Trait, scales = "free", ncol = 2) +
  theme(legend.position="none",
        panel.spacing = unit(0.1, "lines"),
        axis.ticks.x=element_blank()) + theme_bw()
#dev.off()
```


# Joint BLUPS from Disease and Yield traits

```{r joint Blups object of Disease and yield traits, fig.height = 4, fig.width = 3}
AllBlups <- BLUPSDisea %>% left_join(BlupsTraits, by = c("CLONE" = "id")) %>% 
  dplyr::rename(Vigor = Vigor12M)

saveRDS(AllBlups, file = here::here("output", "BLUPsDiseaseAgro.rds"))
```

#### Table 7. BLUPs plus intercept for all disease and agronomic traits for cassava.

```{r table7, echo = FALSE}
library(reactable)
AllBlups %>%
  reactable(columns = list(Anth = colDef(format = colFormat(digits = 2, locales = "en-US")),
                           BlLS = colDef(format = colFormat(digits = 2, locales = "en-US")),
                           BrLS = colDef(format = colFormat(digits = 2, locales = "en-US")),
                           WhLS = colDef(format = colFormat(digits = 2, locales = "en-US")),
                           Vigor = colDef(format = colFormat(digits = 2, locales = "en-US")),
                           NR = colDef(format = colFormat(digits = 2, locales = "en-US")),
                           DRY = colDef(format = colFormat(digits = 2, locales = "en-US")),
                           DMC = colDef(format = colFormat(digits = 2, locales = "en-US")),
                           PTR = colDef(format = colFormat(digits = 2, locales = "en-US")),
                           PPA = colDef(format = colFormat(digits = 2, locales = "en-US"))))
```


[Next page](PCA_DAPC.html)

[Last page](AnalisesDescritivas.html)

[Back to home](index.html)
