---
title: "Home"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

## Introduction

### Phenotypic data
Descriptive and deviance analysis were performed to verify the distribution and the genotypic effects for resistance to foliar diseases.

#### Foliar Disease data
```{r PhenoData Foliar Disease read}
suppressMessages(library(tidyverse))
library(here)
suppressMessages(library(reactable))

# Data read by R using here package to allow the read in any computer with different directory path
PhenoData <- readRDS(here::here("data", "DadosFenotipicos.rds"))

#write.table(PhenoData, file = "data/xx.csv", quote = F, sep = ",", row.names = F)
#PhenoData2 <- read.table("data/xx.csv", header = T, sep = ",")
#saveRDS(PhenoData2, file = here::here("data", "DadosFenotipicos.rds"))
# Add the control information for the Mixed Models analysis
control <- names(table(PhenoData$accession_name)[table(PhenoData$accession_name) > 30])
PhenoData$control <- ifelse(PhenoData$accession_name %in% control, PhenoData$accession_name, "999") 
PhenoData$new <- ifelse(PhenoData$accession_name %in% control, 0, 1)

# Change the Disease names to english abbreviations
colnames(PhenoData)[7:10] <- c("Anth", "WhLS", "BrLS", "BlLS")
```

#### Table 1. Phenotypic data for resistance to foliar disease evaluated at the Cassava Germplasm Bank of EMBRAPA, data collected in 2021.
```{r table1, echo = FALSE}
PhenoData %>% reactable(defaultPageSize = 10)
```
Anthracnosis (Anth), White Leaf spot (WhLS), Brown Leaf spot (BrLS), Blight Leaf Spot (BlLS).

```{r formating foliar disease data to column, message = FALSE, warning = FALSE, eval = TRUE}
lnfUtl <- colnames(PhenoData)[c(1:4, 11:12)]
PhenoData$talhao_number <- NULL
PhenoData$Idade <- NULL
traits <- colnames(PhenoData)[!colnames(PhenoData) %in% lnfUtl]
PhenoData2 <- PhenoData %>% gather(key = traits, value = Y, -all_of(lnfUtl))

saveRDS(PhenoData2, here::here("output", "DadosFenotipicosv2.RDS"))
```

#### Table 2. Data entry for the descriptive analysis and mixed models for resistance of foliar disease in cassava.
```{r table2, echo = FALSE}
PhenoData2 %>% reactable(defaultPageSize = 10)
```
Anthracnosis (Anth), White Leaf spot (WhLS), Brown Leaf spot (BrLS), Blight Leaf Spot (BlLS).

#### Yield Traits
```{r PhenoData Yield traits read}
library(reshape2); library(tidyverse)
DadosProdutivos <- read.table(here::here("data", "Dados Produtivos 2011-2021.CSV"), header = T, sep = ",",
                              na.strings = "NA")

head(DadosProdutivos)

colnames(DadosProdutivos)[c(5, 13:25)] <- c("Genotipo_BGM","AP", "APsF","DMC", "DRY",
                                            "PTR","PRC", "PRNC","PPA","NR","Vigor","Vigor45D","Vigor6M","Vigor12M")

DadosPhen2 <- DadosProdutivos %>% dplyr::select(-c(AP, APsF, PRC, PRNC, Vigor, Vigor45D, Vigor6M))

head(DadosPhen2)
```

```{r formating yield traits data to column}
DadosPhenfin <- reshape2::melt(data = DadosPhen2,id.vars = c("Ano", "Campo",
                                                             "Fazenda", "Local",
                                                             "Delineamento","Controle",
                                                             "Genotipo_BGM","Genotipos",
                                                             "Bloco", "Linha",
                                                             "Coluna", "Stand"),
                                     variable.name = "Trait",
                                     value.name = "Value") %>%
  filter(!is.na(Value)) %>% 
  dplyr::mutate(Ano = Ano,
         Campo = Campo,
         Local = Local,
         trial = match(paste(Ano, Campo, Local, sep = "."), unique(paste(Ano, Campo, Local, sep = "."))),
         studyDesign = Delineamento,
         clone = Genotipo_BGM,
         rep = Bloco,
         check = ifelse(Controle == "1", clone, "999"),
         check = ifelse(clone %in% unique(check), clone, "999"),
         new = ifelse(check != "999", 0, 1),
         y = Value, .keep = "unused") %>% dplyr::select(-c("Genotipos")) %>%
  filter(Trait != "NR" | (Trait=="NR" & y < 30))
head(DadosPhenfin)
saveRDS(object = DadosPhenfin, file = here::here("output", "DadosFenotipicos.rds"))
```


- [Descriptive Analysis](AnalisesDescritivas.html)

### [Mixed Models Analysis](AnalisesModelosMistos.html)

### [Discriminant Analysis of Principal Components](PCA_DAPC.html)

### [Dendrogram and Shannon-Weaver Index](Den_IndSH.html)
