---
title: "Home"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

## Introduction

### Phenotypic data
Descriptive and deviance analysis were performed to verify the distribution and the genotypic effects for resistance to foliar diseases.
In this work we used a natural logarithm transformation for Affected area by Anthracnosis at the innoculated area at stem in cassava plants (AnAr).

```{r}
suppressMessages(library(tidyverse))
library(here)
suppressMessages(library(reactable))

# Data read by R using here package to allow the read in any computer with different directory path
PhenoData <- readRDS(here::here("data", "DadosFenotipicos.RDS"))
# Add the control information for the Mixed Models analysis
control <- names(table(PhenoData$accession_name)[table(PhenoData$accession_name) > 30])
PhenoData$control <- ifelse(PhenoData$accession_name %in% control, PhenoData$accession_name, "999") 
PhenoData$new <- ifelse(PhenoData$accession_name %in% control, 0, 1)

# Change the Disease names to english abbreviations
colnames(PhenoData)[7:11] <- c("Anth", "WhLS", "BrLS", "BlLS", "AnAr")

# Natural logarithm transformation for Anthracnosis Area
PhenoData$AnAr <- log(PhenoData$AnAr)
```

#### Table 1. Phenotypic data for resistance to foliar disease evaluated at the Cassava Germplasm Bank of EMBRAPA, data collected in 2021.
```{r, echo = FALSE}
PhenoData %>% reactable(defaultPageSize = 10)
```
Anthracnosis (Anth), White Leaf spot (WhLS), Brown Leaf spot (BrLS), Blight Leaf Spot (BlLS), Affected area by Anthracnosis at the innoculated area at stem in cassava plants (AnAr).

```{r, message = FALSE, warning = FALSE, eval = TRUE}
lnfUtl <- colnames(PhenoData)[c(1:4, 12:13)]
PhenoData$talhao_number <- NULL
PhenoData$Idade <- NULL
traits <- colnames(PhenoData)[!colnames(PhenoData) %in% lnfUtl]
PhenoData2 <- PhenoData %>% gather(key = traits, value = Y, -all_of(lnfUtl))
```

#### Table 2. Data entry for the descriptive analysis and mixed models for resistance of foliar disease in cassava.
```{r, echo = FALSE}
PhenoData2 %>% reactable(defaultPageSize = 10)
```
Anthracnosis (Anth), White Leaf spot (WhLS), Brown Leaf spot (BrLS), Blight Leaf Spot (BlLS), Affected area by Anthracnosis at the innoculated area at stem in cassava plants (AnAr).

```{r}
saveRDS(PhenoData2, here::here("output", "DadosFenotipicosv2.RDS"))
```

- [Descriptive Analysis](AnalisesDescritivas.html)

### [Mixed Models Analysis](AnalisesModelosMistos.html)

### [Discriminant Analysis of Principal Components](PCA_DAPC.html)

### [Dendrogram and Shannon-Weaver Index](Den_IndSH.html)
